---
title: "Vignette Title"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6, 
  fig.height = 4
)
```

## Introduction

`alocv` is a package that implements the method proposed by Wang *et al*. (2018). In short, it approximates the leave-one-out cross-validation (LOOCV) risk for several standard models, using only the full-data solution. Since the model no longer needs to be fitted $n$ times, the LOOCV estimates can be obtained very efficiently, even in the case when $n$ is so large that leave-one-out is infeasible.

This package currently supports LOOCV approximation for Elastic-net GLM (`glmnet`) and SVM (`e1071`).

## Usage with `glmnet`
We begin by loading our package, of course:
```{r, eval = TRUE, echo = TRUE}
library(alocv)
library(glmnet)
```

For demonstration, we will now create some artifical data of size $n=500$ and $p=200$, with $k=100$ non-zero coefficients:
```{r, eval = TRUE}
n = 500
p = 200
k = 100
beta = rnorm(p, 0, 1)
beta[-(1:k)] = 0

X = matrix(rnorm(n * p, 0, sqrt(1 / k)), n, p)
y = X %*% beta + rnorm(n, 0, 0.5)
y[y >= 0] = 2 * sqrt(y[y >= 0])
y[y < 0] = -2 * sqrt(-y[y < 0])
```

For such a dataset, even the `glmnet` with its state-of-art implementation needs some time to compute the LOOCV estimate:
```{r, eval = TRUE}
ptm = proc.time()
CV_el = cv.glmnet(X, y, alpha = 0.5, nfolds = n, 
                  grouped = F, intercept = T, standardize = T, type.measure = "mse")
proc.time() - ptm
```
In comparison, `alocv` requires only fractions of a second to produce an LOOCV estimation:


```{r, eval = TRUE}
ptm = proc.time()
GLM_el = glmnet(X, y, lambda = CV_el$lambda, alpha = 0.5, intercept = T, standardize = T)
ALO_el = alo_glmnet(X, y, alpha=0.5, standardize=T, intercept=T)
proc.time() - ptm
```
Not only that, the estimation is very accurate, as compared to the true LOOCV risk:

```{r}
plot(CV_el$cvm, xlab = "lambda", ylab = "Risk", type = "l", lwd = 2, col = "darkorange")
lines(ALO_el$alo, type = "b", pch = 4, lwd = 2, col = 4)
```

## Usage with `e1071`

## Reference
K. Rahnama Rad, A. Maleki (2018). [*A scalable estimate of the extra-sample prediction error via approximate leave-one-out*](https://arxiv.org/abs/1801.10243).

S. Wang, W. Zhou, H. Lu, A. Maleki, and V. Mirrokni (2018). [*Approximate Leave-One-Out Approximatet for Fast Parameter Tuning in High Dimensions*](https://arxiv.org/abs/1807.02694).
